\documentclass{tufte-handout}
\usepackage{graphicx}
\usepackage{listings}
\graphicspath{ {./img} }
\title{Synchronzation Examples}
\author{Andr\'es Ponce}

\begin{document}
\maketitle
\begin{abstract}

There are some other problems with synchronization, sucha as the critical section problem
already discussed.
\end{abstract}

Remember that we had discussed \textbf{semaphores}, which were either binary or integer 
variables used to indicate the availability of a critical section and whether a process could
access such critical section. In reality, those could all act as a mutex lock rather than a 
strict binary semaphore.

\subsection{Bounded Buffers Problem}
Remember this problem refers to the change of some variable by two different processes.

The essence of this problem is that we have a \textbf{consumer} and a \textbf{producer}.
The former program is in charge of producing content to go into one of $n$ buffers.
\begin{lstlisting}
int n;
semaphore mutex = 1;
semaphore empty = n;
semaphore full = 0;
\end{lstlisting}

The two processes share the above data structures. The problem lies in how to coordinate
them so that no data is lost or copied. We should thus avoid having the producer 
write to the buffer when it is full, and we should also avoid having the consumer
consume from an empty buffer.

\subsection{Readers-Writers Problem}
This problem can be illustrated with a database. If multiple processes want to read 
from the database, then there is no issue, these are the \textbf{reader} programs.
\textbf{Writer} programs might take a look at the database and try to change some of 
the values in it. As many reader programs can read the database without any issue, 
but when there is a writer program involved then we need to be careful. When a program 
writes to the database, we need it to have exclusive access to the database.

There are two variations of this problem: one where we allow readers to read 
concurrently, and exclude one of the writers programs. Another solution is to give
priority to the writer programs so that they write as soon as possible. Both of these
solutions result in some amount of starvation, in the first case the writer is left
without access and in the second one the reader is left without access.

A solution might be to keep a \texttt{rw\_mutex} semaphore, which both readers and 
writers have to acquire to change the contents of the database. Multiple readers can
access this lock, but only one writer program at a time may access it. This approach might
be useful in situations where there are more readers than writers, since it would minimize
the amount of writers starved and maximize the concurrent readers accessing the data.

\subsection{Dining Philosophers Problem}
Multiple philosophers gathered around a dinner table want to access the bowl of rice.
When a philosopher wants to eat rice, she grabs the chopstick and starts eating. 
A philospher has to grab one chopstick to each of her sides, so she has to wait until
the philosopher near to her have put down the chopstick before eating. 

The problem here is how do we allocate finite resources among various competing processes
without deadlock.

The \textbf{semaphore} solution involves representing each chopsitck (the contested resource)
, with a semaphore. Thus, when a philosopher wants to pick up her chopstick, it waits
for the other one to become available. However, this could result in all chopsticks being
picked up at the same time and thus being unavailable. If all philosophers grabbed their
left chopstick at once, then all would starve, since they would all infiintely wait
for the other chopstick to become available. We could limit the number of philosophers
that are available to eat, or ensure that a philosopher only picks up chopsticks
when they're both available.

The \textbf{monitor} solution presents a deadlock-free solution but only if we pick 
up the chopsticks if both are available.
\section{Synchronization within the Kernel}

\subsection{Synching in Windows}
Windows provdes a spinlock solution when there is some thread waiting to access
a global resource. THis is becuase spinlocks might be more efficient for smaller 
code segments.\footnote{Remember in a spinlock, the waiting program loops until
it can access the critical section.}

In Linux, we can use synchronization by performing some atomic operations. This means
that during these sorts of operations the processes will not be pre-empted and 
interruped by other processes. However, this is mostly seen in counters. More 
sophisticated locking mechanisms might be required for other scenarios.j

\section{POSIX Synchronization}
The POSIX standard uses mutex locks, semaphores, and events to protect ciritical 
sections of code. Semaphores are just some binary variables that indicate the 
tatus of some resource.

Pthreads are the main way for the POSIX standard by which we use semaphores.
Since we use mutex locks in C to acquire the right to modify the data in some 
programs, we usually associate some conditions with it. These conditions
are called condition variables. If the condition is not true, we can place the
variable in a loop and check it periodically.

\section{Alternative Approaches}
Recently there has been an increasing need for mutli-threaded applications but 
designing such applications that are entirely free from race conditions remains
a tad hard.

\subsection{Transactional Memory}
Coming straight from database design, there is an approach to memory called 
\textbf{transactional memory}. A transaction is a sequence of atomic operations 
that are made and then committed once they are completed. Since operations 
on memory are made atomically, then there cannot be any deadlock.

In this approach, the TM system will automatically take care of seeing which 
oeprations can be handled concurrently, such as multiple read operations.
It is then on the hardware implementation to find how to properly schedule 
threads so that there is no collision.

\subsection{OpenMP}
Recall that OpenMP is a library where we can schedule things to run concurrently. 
We can specify some compiler directives a piece of code that is to be executed
by a number of parallel threads. The \texttt{\#pragma omp critical} directive 
indicates that the following piece of code is to be only executed by a single thread.
This is to inform the rest of the system to only allow one thread to execute 
at the same time. This directive then functions more or less like a mutex lock,
so it would still be up to the programmer to handle possible race conditions.

\subsection{Functional Programming Languages}
Imperative languages such as C, Java, and Python are so becuase they maintain 
a  state of all variables and so during sequential execution of instructions,
the order of the instructions is crucial.

\end{document}
